Embeddings and Attention in Predictive Modeling

### 2022-04-11 manuscript/manuscript.Rmd
Spelling corrections

For Model 4, the neural network with multi-dimensional embeddings, we would need to treat the learned values before we can plot them. Specifically, we need to reduce the dimension of the data to 2D for embeddings with more than two dimensions. The standard approaches for this are principal component analysis (PCA; @shlens2014tutorial) and t-distributed stochastic neighbor embedding (t-SNE; @Maaten2008a), both of which we illustrate. For a technical review of these methods, refer to Appendix \\ref{dim-red}.

@@

| Model                                                                             | RMSE             | MAE           |
|-----------------------------------------------------------------------------------|------------------|---------------|
|  Model 1: GLM (gamma/log link)                                                    |          65 676  |       38 431  |
|  Model 2: MLP with 1-dimensional   embeddings                                     |          60 214  |       37 392  |
|  Model 3: GLM (gamma/log link), categorical   variables replaced with embeddings  |          63 733  |       37 987  |
|  Model 4: MLP with multidimensional   embeddings                                  |          60 772  |       37 587  |
|  Linear regression, predictions capped below   at 0.01                            |          61 285  |       38 962  |

@@

| Model                                                                            | RMSE             | MAE              |
|----------------------------------------------------------------------------------|------------------|------------------|
| Model 2: MLP with 1-dimensional   embeddings                                     |          60 214  |          37 392  |
| Model 5: Simple Attention                                                        |          60 367  |          37 527  |
| Model 6: TabTransformer                                                          |          59 908  |          36 501  |
| Model 7: TabNet                                                                  |          63 619  |          39 668  |
| Linear regression, predictions capped below at 0.01                              |          61 285  |          38 962  |

@@

|Embedding Type       |  RMSE|   MAE|
|:--------------------|------:|------:|
|Embedding            | 68 444| 41 557|
|Contextual embedding | 67 905| 40 559|

@@

# Finetuning the Neural Network Models

@@

|                | ReLU/8 | ReLu/16 | ReLu/32 | ReLu/64 | ReLu/128 | tanh/8 | tanh/16 | tanh/32 | tanh/64 | tanh/128 |
|----------------|--------|---------|---------|---------|----------|--------|---------|---------|---------|----------|
| Model 2 - RMSE | 60 988 | 60 561  | 60 125  | 60 147  | 59 918   | 61 728 | 60 506  | 60 130  | 59 893  | 60 065   |
| Model 4 - RMSE | 60 535 | 60 724  | 60 311  | 60 191  | 60 114   | 60 521 | 60 658  | 60 279  | 59 892  | 59 987   |
| Model 2 - MAE  | 36 604 | 36 791  | 36 936  | 37 140  | 36 816   | 37 206 | 36 759  | 36 921  | 37 103  | 37 177   |
| Model 4 - MAE  | 36 828 | 36 824  | 36 780  | 37 054  | 36 672   | 36 528 | 36 725  | 36 483  | 36 766  | 36 394   |
: Results of fine-tuning models in Section 4.

@@

\textcolor{blue}{Models 5 and 6 add an attention-based mechanism to derive contextual embeddings to the models in Section 4. To find optimal hyperparameters, the best performing variation of Model 4 (tanh/64) was taken as a baseline, and the attention mechanisms described for Models 5 and 6 were then added to this baseline model. It was found that the performance of Model 6 improved as a result of adding the Tab Transformer attention mechanism, and, thus, no further fintetuning was performed on this model. However, despite several attempts at optimizing the hyperparameters of Models 5 (which included modifying the size of the embeddings and the amount of dropout applied), significant improvements beyond those shown in the table below were not achieved.}

@@

|                | Scores |
|----------------|--------|
| Model 5 - RMSE | 60 382 |
| Model 5 - MAE  | 37 426 |
| Model 6 - RMSE | 59 821 |
| Model 6 - MAE  | 36 474 |
: Results of fine-tuning models in Section 5.

>>>>

For Model 4, the neural network with multi-dimensional embeddings, we would need to treat the learned values before we can plot them. Specifically, we need to reduce the dimension of the data to 2D for embeddings with more than two dimensions. The standard approaches for this are principal component analysis (PCA; @shlens2014tutorial) and t-distributed stochastic neighbor embedding (t-SNE; @Maaten2008a), both of which we illustrate. For a technical review of these methods, refer to Appendix \ref{dim-red}.

@@

| Model                                                                             | RMSE             | MAE           |
|-----------------------------------------------------------------------------------|------------------|---------------|
|  Model 1: GLM (gamma/log link)                                                    |          65,676  |       38,431  |
|  Model 2: MLP with 1-dimensional   embeddings                                     |          60,214  |       37,392  |
|  Model 3: GLM (gamma/log link), categorical   variables replaced with embeddings  |          63,733  |       37,987  |
|  Model 4: MLP with multidimensional   embeddings                                  |          60,772  |       37,587  |
|  Linear regression, predictions capped below   at 0.01                            |          61,285  |       38,962  |

@@

| Model                                                                            | RMSE             | MAE              |
|----------------------------------------------------------------------------------|------------------|------------------|
| Model 2: MLP with 1-dimensional   embeddings                                     |          60,214  |          37,392  |
| Model 5: Simple Attention                                                        |          60,367  |          37,527  |
| Model 6: TabTransformer                                                          |          59,908  |          36,501  |
| Model 7: TabNet                                                                  |          63,619  |          39,668  |
| Linear regression, predictions capped below at 0.01                              |          61,285  |          38,962  |

@@

|Embedding Type       |  RMSE|   MAE|
|:--------------------|------:|------:|
|Embedding            | 68,444| 41,557|
|Contextual embedding | 67,905| 40,559|

@@

# Fine-tuning the Neural Network Models

@@

|                | ReLU/8 | ReLu/16 | ReLu/32 | ReLu/64 | ReLu/128 | tanh/8 | tanh/16 | tanh/32 | tanh/64 | tanh/128 |
|----------------|--------|---------|---------|---------|----------|--------|---------|---------|---------|----------|
| Model 2 - RMSE | 60,988 | 60,561  | 60,125  | 60,147  | 59,918   | 61,728 | 60,506  | 60,130  | 59,893  | 60,065   |
| Model 4 - RMSE | 60,535 | 60,724  | 60,311  | 60,191  | 60,114   | 60,521 | 60,658  | 60,279  | 59,892  | 59,987   |
| Model 2 - MAE  | 36,604 | 36,791  | 36,936  | 37,140  | 36,816   | 37,206 | 36,759  | 36,921  | 37,103  | 37,177   |
| Model 4 - MAE  | 36,828 | 36,824  | 36,780  | 37,054  | 36,672   | 36,528 | 36,725  | 36,483  | 36,766  | 36,394   |
: Results of fine-tuning models in Section 4.

@@

\textcolor{blue}{Models 5 and 6 add an attention-based mechanism to derive contextual embeddings to the models in Section 4. To find optimal hyperparameters, the best performing variation of Model 4 (tanh/64) was taken as a baseline, and the attention mechanisms described for Models 5 and 6 were then added to this baseline model. It was found that the performance of Model 6 improved as a result of adding the TabTransformer attention mechanism, and, thus, no further fintetuning was performed on this model. However, despite several attempts at optimizing the hyperparameters of Models 5 (which included modifying the size of the embeddings and the amount of dropout applied), significant improvements beyond those shown in the table below were not achieved.}

@@

|                | Scores |
|----------------|--------|
| Model 5 - RMSE | 60,382 |
| Model 5 - MAE  | 37,426 |
| Model 6 - RMSE | 59,821 |
| Model 6 - MAE  | 36,474 |
: Results of fine-tuning models in Section 5.
